Changes From Preregistration

### 2021-07-14 preprint.Rmd
- implemented major changes after feedback from reviewers

---
title             : "Likes or Dislikes, Gratifications or Concerns? Analyzing Popularity Cues and Privacy Calculus when Communicating Online"
shorttitle        : "Popularity cues & privacy calculus"
abstract: |
    According to privacy calculus, both privacy concerns and expected gratifications explain why people communicate online. So far, most findings were based on self-reports or short-term experiments, and it is still largely unknown how popularity cues such as like and dislike buttons affect privacy calculus. To answer these questions, a preregistered one-week field experiment was conducted. Participants were randomly distributed to three different websites, on which they discussed a current political topic. The websites featured either (a) like buttons, (b) like and dislike buttons, or (c) no like or dislike buttons, and were otherwise identical. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. Although the originally preregistered model was largely rejected, the results showed that a considerable share of actual communication could be explained by privacy concerns, gratifications, privacy deliberation, trust, and self-efficacy. The impact of the popularity cues on privacy calculus and communication was negligible.
keywords          : "privacy calculus, communication, popularity cues, field experiment, structural equation modeling, preregistration"
bibliography      : 
  - [bibliography/r-references.bib]
  - [bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
# to compile as pdf
# (a) move file "_site.yml" to folder "archive"
# (b) uncomment the following lines
output            : papaja::apa6_pdf
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  - \note{\clearpage}
# output            : papaja::apa6_docx
# output:
#   html_document:
#     toc: TRUE
#     toc_float:
#       collapsed: FALSE
#     toc_depth: 2
---

@@

# Introduction

Understanding why people share personal information online is a critical question for society and research. 
Originally, it was assumed that the online sharing of information is erratic and that it cannot be predicted by people's personal beliefs, concerns, or attitudes. 
Most prominently, the privacy paradox stated that people communicate vast amounts of personal information online *despite* having substantial concerns about their privacy [@barnesPrivacyParadoxSocial2006; @taddickenUsesPrivacyOnline2011].

Somewhat surprisingly, and despite its popularity in the media [@newyorkpublicradioPrivacyParadox2018], empirical support for the privacy paradox is ambivalent.  
A recent meta-analysis reported a correlation between privacy concerns and self-disclosure on SNS of *r* = -.13 [@baruhOnlinePrivacyConcerns2017], which shows that privacy concerns are indeed related to communication online.

Rather than further pursuing the privacy paradox, a large share of current day research builds on the so-called _privacy-calculus_ [@lauferPrivacyConceptSocial1977].
The privacy calculus states that communication online can be explained---at least partly---by means of expected risks *and* expected benefits [@krasnovaOnlineSocialNetworks2010].
By operationalizing expected risks as privacy concerns, several studies have shown that experiencing privacy concerns is related to sharing less information online, whereas expecting benefits is related to sharing more information online [@heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

First, current research on the privacy calculus is often criticized for not explicitly focusing on the _deliberation process_ when communicating online. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that both concerns and gratifications correlate with communication behavior online is not sufficient evidence for an explicit weighing process. 
This study, therefore, explicitly focuses on the privacy deliberation process.

Second, in this study I approach the privacy calculus from a theoretical perspective of _bounded rationality_. 
It is likely that other factors next to risks and benefits also determine behavior.
I therefore extend the privacy calculus model theoretically by investigating the role and interplay of trust and self-efficacy.

Third, the privacy calculus does not take place in a vacuum. 
It is often argued that communication online can be easily triggered by external circumstances.
I therefore analyze whether the privacy calculus is affected by the affordances of a website. 
Specifically, I investigate whether *popularity cues* such as like and dislike buttons affect the privacy calculus and whether they foster communication online.

Fourth, it is still largely unknown whether the privacy calculus can be replicated with _behavioral data_ in an authentic long-term setting [@kokolakisPrivacyAttitudesPrivacy2017]. 
Thus far, much research on the privacy calculus used self-reports of behavior [@krasnovaOnlineSocialNetworks2010], vignette approaches [@bolUnderstandingEffectsPersonalization2018], or one-shot experiments in the lab [@treptePrivacyCalculusContextualized2020]. 
A long-term field study observing actual behavior in an authentic context is still missing.

To test the research questions, a representative sample of the German population was collected in a preregistered online field experiment.
Participants were randomly distributed to one of three different websites, which either included a like button, both a like and a dislike button, or no buttons at all. 
Over the course of one week, participants had the chance to discuss a topical issue (i.e., prevention of terrorist attacks in Germany). 
Afterward, they answered a follow-up questionnaire with items measuring the privacy calculus variables.

@@

The key variable of interest for this study is (verbal) communication online.
Are people willing to engage in a conversation?
Do they express their opinion? 
In communicating online, people share much information about themselves.
Communication is, hence, closely related to self-disclosure, and it is a primary means of regulating privacy [e.g., @dienlinPrivacyProcessModel2014].
<!-- Almost all types of communication allow for meaningful inferences about a person, even when we _don't_ communicate [@watzlawickPragmaticsHumanCommunication2011].  -->
<!-- However, the recent years have illustrated vividly how much additional insights we can get about a person, simply by analyzing their digital traces of communication or their meta-data [@kosinskiPrivateTraitsAttributes2013].  -->
<!-- The more we communicate, the more we self-disclose. -->

Privacy concerns were defined as follows.
"Taken together, concerns about online privacy represent how much an individual is motivated to focus on their control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that their privacy might be threatened" [@dienlinLongitudinalAnalysisPrivacy2021, p. 4].

In this study I adopt the theoretical perspective of the privacy calculus [@lauferPrivacyConceptSocial1977].
The privacy calculus assumes that when communicating online people engage in a rational weighing of risks and benefits.
Notably, I don't assume that this weighing process is flawless or that humans are perfect rational agents.
Instead, I understand the privacy calculus from the perspective of _bounded rationality_ [@simonBoundedRationality1990].
Bounded rationality has three tenets: 
"(1) humans are cognitively constrained; 
(2) these constraints impact decision making; 
and (3) difficult problems reveal the constraints and highlight their significance." [@bendorBoundedRationality2015, p. 1303]
Crucially, although bounded rationally upholds that human behavior is not perfectly logical, this does not meant that it is irrational [@gigerenzerBoundedRationalityAdaptive2002].
Instead, it is a continuum.
Humans are still trying to optimize the outcomes of their behavior according to their own best interests or values.
It is only that their capacity to do so is bounded.

Transferred to the context of online privacy, it is by now well known that several irregularities and inconsistencies between concerns and communication behavior exist.
These differences stem from, for example, information asymmetries, present bias, intangibility, illusory control, or herding [@acquistiSecretsLikesDrive2020].
At the same time, _on average_ people do behave according to their interests, respond to incentives, or actively manage their privacy [@baruhOnlinePrivacyConcerns2017; @dienlinExtendedPrivacyCalculus2016; @soloveMythPrivacyParadox2020].

I therefore hypothesize that people who experience more privacy concerns engage in less communication online.
In light of bounded rationality and the existence of other competing factors that also influence online-communication (see below), the effect is likely small.

In turn, the most relevant factor driving online communication is _expected gratifications_.
People accept a loss of privacy if they can gain something in return [e.g., @lauferPrivacyConceptSocial1977].
The most prominent gratifications of online communication include social support [@krasnovaOnlineSocialNetworks2010], social capital [@ellisonNegotiatingPrivacyConcerns2011], entertainment [@dhirUnderstandingRelationshipIntensity2017], information-seeking [@whitingWhyPeopleUse2013], and self-presentation [@minHowArePeople2015].
Several studies have shown, that gratifications outweigh concerns [@dienlinExtendedPrivacyCalculus2016; @bolUnderstandingEffectsPersonalization2018].
As a result, we expect a moderate relationship.

H1: People who are more concerned about their privacy than others are less likely to communicate actively on a website.

H2: People who obtain more gratifications from using a website are more likely to communicate actively on a website.

Privacy calculus implies that people _explicitly_ compare benefits and disadvantages before communicating online.
Research on the privacy calculus has often ignored this aspect [@knijnenburgDeathPrivacyCalculus2017].
Only observing that privacy concerns or expected gratifications and communication online are _related_ is insufficient to prove an explicit weighing process.
Hence, we here analyze how much people actively deliberate about their privacy and how that might influence the privacy calculus.

We can understand the privacy calculus from two perspectives (Table \@ref(tab:PC)):
First, is the communication behavior aligned with people's privacy concerns and expected benefits? 
Second, is the communication process implicit or explicit?

Here, I suggest that the privacy calculus should be discussed in light of dual process theories, which state that people either deliberately, explicitly, and centrally take decisions, or instead do so automatically, implicitly, and peripherally [@kahnemanThinkingFastSlow2011; @pettyCommunicationPersuasionCentral1986]. 
Accordingly, privacy calculus would assume that people, when it comes to disclosing, engage in a central processing. 
Building on @omarzuDisclosureDecisionModel2000 and @altmanPrivacyConceptualAnalysis1976, I hence introduce and investigate a novel concept termed _privacy deliberation_.
Privacy deliberation captures the extent to which individual people explicitly compare potential positive and negative outcomes before communicating with others.

On the one hand, deliberating about privacy could _reduce_ subsequent communication.
Refraining from communication---the primary means of connecting with others---likely requires some active and deliberate restraint. 
This is especially true for social media, which are designed to elicit communication and participation.
Actively thinking about whether communicating is really worthwhile might be the first step not to participate.
On the other hand, deliberating about privacy might also _increase_ communication.
A person concerned about their privacy might conclude that in this situation communication is actually beneficial. 
Deliberation could represent some kind of inner consent, providing additional affirmation.

Alternatively, it could be that deliberation functions as a moderator.
For example, if people actively deliberate about whether or not to disclose, this might reinforce the effect of concerns or gratifications.
Reflecting about the pros and cons of communication might concerns and gratifications more salient. 
Alternatively, it could also be that deliberating decreases the effects, for example because apparent gratifications are considered more critically, and maybe loose their appeal.

I therefore formulate the following two research questions:

RQ1: Do people who deliberate more actively whether they should communicate, communicate more or less online?

RQ2: Do people who deliberate more actively whether they should communicate, show larger or smaller relations between concerns, gratifications and communication behavior?

Bounded rationality implies that additional factors should also explain communication.
Communication online often takes place in situations where information is limited or obscure.
The more familiar users are with a context, the more experience, knowledge, and literacy they possess, the more likely they should be to navigate online contexts successfully.
In other words, if users possess more _self-efficacy_ to participate, they should also communicate more.
Related, people who report more privacy self-efficacy also engage in more self-withdrawal [@chenRevisitingPrivacyParadox2018; @dienlinExtendedPrivacyCalculus2016].

H3: People are more likely to communicate on a website when their self-efficacy about self-disclosing on the website is higher.

@@

Trust can be conceptualized in two different ways [@gefenTrustTAMOnline2003]. 
It either captures "_specific_ beliefs dealing primarily with the integrity, benevolence, and ability of another party" [@gefenTrustTAMOnline2003, p. 55, emphasis added].
Alternatively, it refers to a "_general_ belief that another party can be trusted" [@gefenTrustTAMOnline2003, p.55, emphasis added]. 
Whereas specific trust focuses on the causes of trust, general trust emphasizes the experience of trust.
<!-- @gefenTrustTAMOnline2003 prioritize specific trust (p. 60).  -->
In the online context, there exist several different _targets_ of trust, including (a) the information system, (b) the provider, (c) the Internet, and (d) the community of other users [@sollnerWhyDifferentTrust2016]. 
Because the targets can be largely different, it is often recommended to analyze them individually.

H4: People are more likely to communicate on a website when they have greater trust in the provider, the website, and the other users.

@@

So far I analyzed user-oriented factors that explain communication online.
But how does the context, the digital infrastructure, affect the privacy calculus and communication? 
In what follows I do not focus on specific _features_ of particular websites, which can change and quickly become obsolete [@foxDistinguishingTechnologiesSocial2017]. 
Instead, I address the underlying latent structures by analyzing so-called _affordances_ [@ellisonSocialNetworkSite2015; @foxDistinguishingTechnologiesSocial2017]. 
Developed by @gibsonEcologicalApproachVisual2015, affordances emphasize that it is not the _objective features_ of objects that determine behavior, but rather our _subjective perceptions_. 
Affordances are mental representations of how objects might be used.
<!-- ; as such, they are by definition subjective.  -->
There is an ongoing debate on what exactly defines an affordance [@evansExplicatingAffordancesConceptual2017].
For example, whereas @evansExplicatingAffordancesConceptual2017 propose three affordances for mediated communication (i.e., anonymity, persistence, and visibility), @foxDistinguishingTechnologiesSocial2017 suggest 10 affordances for SNSs alone (i.e., accessibility, bandwidth, social presence, privacy, network association, personalization, persistence, editability, conversation control, and anonymity).

The privacy calculus states that both benefits and costs determine behavior.
Popularity cues such as like and dislike buttons, which are categorized as "paralinguistic digital affordances" [@carrPredictingThresholdPerceived2018, p. 142], can be linked to the two sides of the privacy calculus. 
The like button is positive and a potential benefit:
It expresses an endorsement, a compliment, a reward [@carrPredictingThresholdPerceived2018; @sumnerFunctionalApproachFacebook2017]. 
The dislike button is negative and a potential cost: 
It expresses criticism and a way to downgrade content.

Paralinguistic digital affordances and specifically popularity cues can affect behavior [@kramerMasteringChallengeBalancing2020; @treptePrivacyCalculusContextualized2020].
Online comments that already have several dislikes are much more likely to receive further dislikes [@muchnikSocialInfluenceBias2013]. 
When users disagree with a post, they are more likely to click on a button labeled _respect_ compared to a button labeled _like_ [@stroudRecommendRespectAltering2017].
The potentially stark negative effect of the dislike button might also explain why to date only a handful of major websites have implemented it (e.g., youtube, reddit, or stackexchange).
In this vein, popularity cues likely also impact the privacy calculus [@kramerMasteringChallengeBalancing2020].

@@

Receiving a _dislike_ should feel more like a punishment.
Dislikes introduce a _loss frame_. 
<!-- Although most communication emphasizes positive aspects, the Internet is also replete with spite, envy, and arguments. -->
As a result, websites featuring both like _and_ dislike buttons should be more ambivalent compared to websites without any popularity cues.
In online contexts, gains often outweigh losses.
Having both types of popularity cues might still lead to more gratifications and communication.
However, privacy concerns should not be reduced anymore:
People who are more concerned about their privacy are also more shy and risk averse [@dienlinPsychologyPrivacyAnalyzing2017].
Implementing the dislike button might therefore increase privacy concerns, thereby canceling out the positive effects of the like button. 
And because there is more at stake, participants should deliberate even more whether or not to disclose.

There are two potential underlying theoretical pathways:
The _mere presence_ of popularity cues might affect whether people are willing to disclose; being able to attract likes might motivate users to communicate, while the mere option to receive dislikes might intimidate others.
On the other hand, _actually receiving_ likes or dislikes might then affect subsequent behavior, potentially reinforcing the process.

H5.	Compared to people who use a website without like or dislike buttons, people who use a website with like buttons (a) communicate more, (b) obtain more gratifications, (c) are less concerned about their privacy, and (d) deliberate more about whether they should communicate online.

H6.	Compared to people who use a website without like or dislike buttons, people who use a website with like _and_ dislike buttons (a) communicate more, (b) obtain more gratifications, and (c) deliberate more about whether they should communicate online.

<!-- When directly comparing websites including both like and dislike buttons with websites including only like buttons, building on the rationales presented above, it is likely that websites including both buttons should increase privacy concerns and privacy deliberation. -->

@@

For a simplified overview of the analyzed model, see Figure \@ref(fig:model).

@@

The online supplementary material (OSM) of this study includes the data, research materials, analyses scripts, and a reproducible version of this manuscript, which can be found on the manuscript's companion website (https://XMtRa.github.io/privacy_calc_exp_anon).
I preregistered the study using the registration form _OSF Prereg_, which includes the hypotheses, sample size, research materials, analyses, and exclusion criteria (see https://osf.io/a6tzc/?view_only=5d0ef9fe5e1745878cd1b19273cdf859). 
I needed to change the pre-defined plan in some cases. 
For a full account of all changes, see OSM. 
New analyses that were not preregistered appear in the section Exploratory Analyses.

@@

I collaborated with a market research company to recruit participants. 
As incentive, participants were awarded digital points, which they could use to get special offers from other online commerce services. 
Participants were above the age of 18 and lived in Germany. 
In a first step, the company sent its panel members an invitation to participate in the study (_invitation_). 
In this invitation, panel members were asked to participate in a study analyzing the current threat posed by terrorist attacks in Germany.[^1] 
Members who decided to take part were subsequently sent the first questionnaire (_T1_), in which I (a) asked about their sociodemographics, (b) provided more details about the study, and (c) included a registration link for the website, which was described as "participation platform".
Afterward, participants were randomly assigned to one of the three websites. 
After registration was completed, participants were invited (but not obliged) to discuss the topic of the terrorism threat in Germany over the course of one week (_field_). 
Subsequently, participants received a follow-up questionnaire in which the self-reported measures were collected (_T2_). 
Measures were collected after and not before the field phase in order not to prime participants or reveal the primary research interest.

>>>>

---
title             : "How Do Like and Dislike Buttons Affect Communication? Testing the Privacy Calculus in a Preregistered One-Week Field Experiment"
shorttitle        : "Popularity cues and privacy calculus"
author:            
  - name          : "Dienlin, Tobias"
    email         : "tobias.dienlin@univie.ac.at"
    affiliation   : "1"
    corresponding : yes
    address       : "University of Vienna, Department of Communication, 1090 Vienna, Austria"
  - name          : "Braeunlich, Katharina"
    email         : "braeunlich@uni-koblenz.de"
    affiliation   : "2"
  - name          : "Trepte, Sabine"
    email         : "sabine.trepte@uni-hohenheim.de"
    affiliation   : "3"
affiliation:
  - id            : "1"
    institution   : "University of Vienna"
  - id            : "2"
    institution   : "University of Koblenz-Landau"
  - id            : "3"
    institution   : "University of Hohenheim"
note: |
  This preprint has been submitted to a  journal and is currently under review. Please cite carefully.

authornote: |
  Dr. Tobias Dienlin is Assistant Professor of Interactive Communication at University of Vienna. He received his Ph.D. from University of Hohenheim.

  Dr. Katharina Braeunlich works at the Federal Office for Information Security in Bonn. She is an alumna from University of Koblenz-Landau, where she received her Ph.D.

  Dr. Sabine Trepte is Full Professor for Media Psychology at University of Hohenheim. She received her Ph.D. from Hanover University of Music, Drama and Media.

  All authors contributed extensively to the work presented in this paper. TD, KB, & ST designed the study; KB & TD designed the online website; TD & KB administered the data collection and importation; TD wrote the code, ran the models, and analyzed the output data; TD wrote the manuscript and ST provided comments; ST supervised the project.

    The authors declare no competing interests.

    This research was funded by the Volkswagen Foundation, project "Transformations of privacy", which was awarded to Sandra Seubert, Sabine Trepte, Ruediger Grimm, & Christoph Gusy. We would like to thank all our colleagues from the project as well as Niklas Johannes for valuable feedback.

    This manuscript features a companion website, which includes the data, code, additional analyses, the preregistration, and a reproducible version of the manuscript (https://XMtRa.github.io/privacy_calc_exp_anon).
abstract: |
    According to privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, most findings were based on self-reports or short-term experiments, and it is still largely unknown how popularity cues such as like and dislike buttons affect privacy calculus. To answer these questions we ran a preregistered one-week field experiment. Participants were randomly distributed to three different websites, on which they discussed a current political topic. The websites featured either (a) like buttons, (b) like and dislike buttons, or (c) no like or dislike buttons, and were otherwise identical. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. Although the originally preregistered model was rejected, the results showed that a considerable share of actual self-disclosure could be explained by privacy concerns, gratifications, privacy deliberation, trust, and self-efficacy. The impact of the popularity cues on self-disclosure and privacy calculus was negligible.
keywords          : "privacy calculus, self-disclosure, popularity cues, field experiment, structural equation modeling, preregistration"
bibliography      : 
  - [bibliography/r-references.bib]
  - [bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
wordcount         : "`r wordcountaddin::word_count('manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
# to compile as pdf
# (a) move file "_site.yml" to folder "archive"
# (b) uncomment the following lines
output            : papaja::apa6_pdf
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  - \note{\clearpage}
# output            : papaja::apa6_docx
# output:
#   html_document:
#     toc: TRUE
#     toc_float:
#       collapsed: FALSE
#     toc_depth: 2
---

@@

<!-- # Introduction -->

Understanding why people disclose personal information online is a critical question for society and research. 
Originally, it was assumed that online self-disclosure is erratic and that it cannot be predicted by people's personal beliefs, concerns, or attitudes. 
Most prominently, the privacy paradox stated that people self-disclose vast amounts of personal information online *despite* having substantial concerns about their privacy [@barnesPrivacyParadoxSocial2006; @taddickenUsesPrivacyOnline2011].

Somewhat surprisingly, and despite its popularity in the media [@newyorkpublicradioPrivacyParadox2018], the privacy paradox has garnered comparatively little empirical support. 
A recent meta-analysis reported a correlation between privacy concerns and self-disclosure on SNS of *r* = -.13 [@baruhOnlinePrivacyConcerns2017], which shows that privacy concerns are indeed related to self-disclosure online.

Hence, rather than further pursuing the privacy paradox, a large share of current day research builds on the so-called _privacy-calculus_ [@lauferPrivacyConceptSocial1977].
The privacy calculus states that self-disclosure online can be explained---at least partly---by means of expected risks *and* expected benefits [@krasnovaOnlineSocialNetworks2010].
By operationalizing expected risks as privacy concerns, several studies have shown that experiencing privacy concerns is related to disclosing less information online, whereas expecting benefits is related to disclosing more information online [@heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

First, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process behind self-disclosure. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that both concerns and gratifications correlate with self-disclosure is not sufficient evidence for an explicit weighing process. 
In this study, we therefore explicitly focus on the privacy deliberation process.

Second, we approach the privacy calculus from a theoretical perspective of bounded rationality. 
It is likely that other factors next to risks and benefits also determine behavior.
We therefore extend the privacy calculus model theoretically by investigating the role and interplay of trust and self-efficacy.

Third, the privacy calculus does not take place in a vacuum, and it is often argued that self-disclosure can be easily triggered by external circumstances.
We therefore analyze whether the privacy calculus is affected by the affordances of a website. 
Specifically, we investigate whether *popularity cues* such as like and dislike buttons affect the privacy calculus and whether they foster self-disclosure.

Fourth, it is still largely unknown whether the privacy calculus can be replicated with behavioral data in an authentic long-term setting [@kokolakisPrivacyAttitudesPrivacy2017]. 
Thus far, most research on the privacy calculus used self-reports of behavior [@krasnovaOnlineSocialNetworks2010], vignette approaches [@bolUnderstandingEffectsPersonalization2018], or one-shot experiments in the lab [@treptePrivacyCalculusContextualized2020]. 
A long-term field study in which actual behavior is observed in an authentic context is still missing.

To test our research questions, we collected a representative sample of the German population and conducted a preregistered online field experiment.
Participants were randomly distributed to one of three different websites, which either included a like button, both a like and a dislike button, or no buttons at all. 
Over the course of one week participants had the chance to discuss a topical issue (i.e., prevention of terrorist attacks in Germany). 
Afterward, they answered a follow-up questionnaire with items measuring the privacy calculus variables.

@@

The key variable of interest for this study is self-disclosure.
Self-disclosure is a primary means of regulating privacy [e.g., @masurSituationalPrivacySelfdisclosure2018].
There are two different understandings of self-disclosure: 
The first limits self-disclosure to _deliberate_ acts of sharing _truthful_ information about the self with others [@jourardTransparentSelf1964]. 
The second considers _all_ acts of communication---active or passive, deliberate or unintended---as self-disclosure.
In other words, when communicating it is not possible not to self-disclose.
In this study, we adopt the latter understanding.
All types of communication allow for meaningful inferences about a person [@watzlawickPragmaticsHumanCommunication2011]. 
The recent years have illustrated vividly how much we can learn about a person only by analyzing their digital traces of communication or their meta-data [@kosinskiPrivateTraitsAttributes2013]. 
Self-disclosure has three different dimensions: breadth (i.e., number of topics covered), depth (i.e., intimacy of topics covered), and length (i.e., quantity of disclosure) [@omarzuDisclosureDecisionModel2000]. 
Here, we focus on _communication quantity_.
The more we communicate, the more we self-disclose.
Notably, the relation is not linear: 
Impressions are formed quickly, and the more we communicate the less likely it becomes to share new information.

Privacy concerns were defined as follows: 
"Concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened" [@dienlinLongitudinalAnalysisPrivacy2019, p. 6].

In this study we adopt the theoretical perspective of the privacy calculus [@lauferPrivacyConceptSocial1977].
The privacy calculus assumes that when self-disclosing people engage in a rational weighing of risks and benefits.
We do not assume that this weighing process is flawless and that humans are perfect rational agents.
Instead, we understand the privacy calculus from the perspective of _bounded rationality_ [@simonBoundedRationality1990].
Although humans weigh pros and cons when disclosing, their capacity is limited, errors happen, and manipulation always possible.

We therefore hypothesize that experiencing privacy concerns should reduce self-disclosure.
In light of bounded rationality and the existence of other competing factors that also influence self-disclosure, we assume that the effect is only small.
Previous research has likewise found that people who are more concerned about their privacy than others are slightly less likely to share personal information [@baruhOnlinePrivacyConcerns2017; @heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

H1: People are more likely to self-disclose on a website when they are less concerned about their privacy.

According to privacy calculus, the most relevant competing factor explaining self-disclosure is _expected gratifications_.
People accept a loss of privacy if they can gain something in return [e.g., @lauferPrivacyConceptSocial1977].
The most prominent gratifications include social support [@krasnovaOnlineSocialNetworks2010], social capital [@ellisonNegotiatingPrivacyConcerns2011], entertainment [@dhirUnderstandingRelationshipIntensity2017], information-seeking [@whitingWhyPeopleUse2013], and self-presentation [@minHowArePeople2015].

H2: People are more likely to self-disclose on a website when they obtain more gratifications from using the website.

Privacy calculus holds that people deliberately compare benefits and disadvantages when disclosing.
However, that process has not been analyzed explicitly so far [@knijnenburgDeathPrivacyCalculus2017].
Only observing that privacy concerns and expected gratifications are related to self-disclosure is by itself not sufficient to prove an explicit weighing process.
Here, we suggest that the privacy calculus should be discussed in light of dual process theories, which state that people either deliberately, explicitly, and centrally take decisions, or instead do so automatically, implicitly, and peripherally [@kahnemanThinkingFastSlow2011; @pettyCommunicationPersuasionCentral1986]. 
Accordingly, privacy calculus would assume that people, when it comes to disclosing, engage in a central processing. 
Building on @omarzuDisclosureDecisionModel2000 and @altmanPrivacyConceptualAnalysis1976, we hence introduce and investigate a novel concept termed _privacy deliberation_.
Privacy deliberation captures the extent to which individual people explicitly compare potential positive and negative outcomes before communicating with others.

On the one hand, deliberating about privacy could _reduce_ subsequent self-disclosure.
Refraining from communication---the primary means of connecting with others---likely requires some active and deliberate restraint. 
On the other hand, deliberating about privacy might also _increase_ self-disclosure.
A person concerned about their privacy might conclude that in this situation self-disclosure is actually beneficial. 
Deliberation could represent some kind of inner consent, providing additional affirmation.
We therefore formulate the following research question:

RQ1: Are people more or less likely to self-disclose on a website depending on how actively they deliberate about whether they should self-disclose?

Bounded rationality implies that additional factors should also explain self-disclosure.
Self-disclosure online often takes place in situations where information is limited or obscure.
The more familiar users are with a context, the more experience, knowledge, and literacy they possess, the more likely they should be to navigate online contexts successfully.
In other words, if users possess more _self-efficacy_ regarding self-disclosing, they should also self-disclose more.
Related, people who report more privacy self-efficacy also engage in more self-withdrawal [@chenRevisitingPrivacyParadox2018; @dienlinExtendedPrivacyCalculus2016].

H3: People are more likely to self-disclose on a website when their self-efficacy about self-disclosing on the website is higher.

@@

Trust can be conceptualized in two different ways [@gefenTrustTAMOnline2003]. 
It either captures "_specific_ beliefs dealing primarily with the integrity, benevolence, and ability of another party" [@gefenTrustTAMOnline2003, p. 55, emphasis added].
Alternatively, it refers to a "_general_ belief that another party can be trusted" [@gefenTrustTAMOnline2003, p.55, emphasis added]. 
Whereas specific trust focuses on the causes of trust, general trust emphasizes the experience of trust.
<!-- @gefenTrustTAMOnline2003 prioritize specific trust (p. 60).  -->
In the online context, there exist several different _targets_ of trust, including (a) the information system, (b) the provider, (c) the Internet, and (d) the community of other users [@sollnerWhyDifferentTrust2016].

H4: People are more likely to self-disclose on a website when they have greater trust in the provider, the website, and the other users.

@@

So far we analyzed user-oriented factors that explain self-disclosure.
But how does the context, the digital infrastructure, affect the privacy calculus and self-disclosure? 
In what follows we do not focus on specific _features_ of particular websites, which can change and quickly become obsolete [@foxDistinguishingTechnologiesSocial2017]. 
Instead, we address the underlying latent structures by analyzing so-called _affordances_ [@ellisonSocialNetworkSite2015; @foxDistinguishingTechnologiesSocial2017]. 
Developed by @gibsonEcologicalApproachVisual2015, affordances emphasize that it is not the _objective features_ of objects that determine behavior, but our _subjective perceptions_. 
Affordances are mental representations of how objects might be used.
<!-- ; as such, they are by definition subjective.  -->
<!-- There is an ongoing debate on what exactly defines an affordance [@evansExplicatingAffordancesConceptual2017]. 
For example, whereas @evansExplicatingAffordancesConceptual2017 propose three affordances for mediated communication (i.e., anonymity, persistence, and visibility), @foxDistinguishingTechnologiesSocial2017 suggest 10 affordances for SNSs alone (i.e., accessibility, bandwidth, social presence, privacy, network association, personalization, persistence, editability, conversation control, and anonymity).  -->

The privacy calculus states that both benefits and costs determine behavior.
Popularity cues such as like and dislike buttons, which are categorized as "paralinguistic digital affordances" [@carrPredictingThresholdPerceived2018, p. 142], can be linked tp the two sides of the privacy calculus. 
The like button is positive and a potential benefit:
It expresses an endorsement, a compliment, a reward [@carrPredictingThresholdPerceived2018; @sumnerFunctionalApproachFacebook2017]. 
The dislike button is negative and a potential cost: 
It expresses criticism and a way to downgrade content.

Paralinguistic digital affordances and specifically popularity cues can affect behavior [@kramerMasteringChallengeBalancing2020; @treptePrivacyCalculusContextualized2020].
Online comments that already have several dislikes are much more likely to receive further dislikes [@muchnikSocialInfluenceBias2013]. 
When users disagree with a post, they are more likely to click on a button labeled _respect_ compared to a button labeled _like_ [@stroudRecommendRespectAltering2017].
<!-- The potentially stark negative effect of the dislike button might also explain why to date only a handful of major websites have implemented it (e.g., youtube, reddit, or stackexchange). -->
In this vein, popularity cues likely also impact the privacy calculus [@kramerMasteringChallengeBalancing2020].

@@

H5.	Compared to people who use a website without like or dislike buttons, people who use a website with like buttons (a) self-disclose more, (b) obtain more gratifications, (c) are less concerned about their privacy, and (d) deliberate more about whether they should communicate online.

Receiving a _dislike_ should feel more like a punishment.
Dislikes introduce a _loss frame_. 
<!-- Although most communication emphasizes positive aspects, the Internet is also replete with spite, envy, and arguments. -->
As a result, websites featuring both like _and_ dislike buttons should be more ambivalent compared to websites without any popularity cues.
In online contexts, gains often outweigh losses.
Having both types of popularity cues might still lead to more gratifications and self-disclosure.
However, privacy concerns should not be reduced anymore:
People who are more concerned about their privacy are also more shy and risk averse [@dienlinPsychologyPrivacyAnalyzing2017].
Implementing the dislike button might therefore increase privacy concerns, thereby canceling out the positive effects of the like button. 
And because there is more at stake, participants should deliberate even more whether or not to disclose.

H6.	Compared to people who use a website without like or dislike buttons, people who use a website with like _and_ dislike buttons (a) self-disclose more, (b) obtain more gratifications, and (c) deliberate more about whether they should communicate online.

When directly comparing websites including both like and dislike buttons with websites including only like buttons, building on the rationales presented above, it is likely that websites including both buttons should increase privacy concerns and privacy deliberation.

@@

For a simplified overview of the model we analyzed, see Figure \@ref(fig:model).

@@

The online supplementary material (OSM) of this study includes the data, research materials, analyses scripts, and a reproducible version of this manuscript, which can be found on the manuscript's companion website (https://XMtRa.github.io/privacy_calc_exp_anon).
We preregistered the study using the registration form _OSF Prereg_, which includes the hypotheses, sample size, research materials, analyses, and exclusion criteria (see https://osf.io/a6tzc/?view_only=5d0ef9fe5e1745878cd1b19273cdf859). 
We needed to change our pre-defined plan in some cases. 
For a full account of all changes, see OSM. 
New analyses that were not preregistered appear in the section Exploratory Analyses.

@@

We collaborated with a market research company to recruit participants. 
As incentive, participants were awarded digital points, which they could use to get special offers from other online commerce services. 
Participants were above the age of 18 and lived in Germany. 
In a first step, the company sent its panel members an invitation to participate in the study (_invitation_). 
In this invitation, panel members were asked to participate in a study analyzing the current threat posed by terrorist attacks in Germany.[^1] 
Members who decided to take part were subsequently sent the first questionnaire (_T1_), in which we (a) asked about their sociodemographics, (b) provided more details about the study, and (c) included a registration link for the website, which was described as "participation platform".
Afterward, participants were randomly assigned to one of the three websites. 
After registration was completed, participants were invited (but not obliged) to discuss the topic of the terrorism threat in Germany over the course of one week (_field_). 
Subsequently, participants received a follow-up questionnaire in which the self-reported measures were collected (_T2_). 
Measures were collected after and not before the field phase in order not to prime participants or reveal our primary research interest.

[^1]: Although the terror attack was not of primary interest for this study, the data can and will also be used to analyze perceptions of the terrorism threat. Hence, no deception took place, and in the debriefing participants were informed about our additional research interest in privacy.
